---
title: "PA1_template"
author: "Parker"
date: "2017年11月13日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction

###This project needs us to load a data set, performs some basic analysis, and produces some charts by following five steps in terms of R markdown document

**1. Loading and preprocessing the data**

After set up working directory into the path where activity dataset is saved

```{r rawdata}
if (! exists("activity")){
    activity<- read.csv("activity.csv")
}
head(activity)
```

```{r total_steps_perday}
data <- aggregate(steps ~ date, data=activity, sum, na.rm = TRUE)
head(data)
hist(data$steps, breaks=40, main="Total Steps per Day", xlab="Steps", ylab="Frequency", col = 3)
```

In conclusion, in terms of histgram, it is roughly seen that median number is around 1000.

**2. What is mean total number of steps taken per day?**

```{r mean_steps_perday}
data_mean <- mean(data$steps, na.rm=TRUE)
data_median <- median(data$steps, na.rm=TRUE)
print(paste("The mean steps per day is: ", data_mean))
print(paste("The median steps per day is: ", data_median))
```

**3. What is the average daily activity pattern?**

```{r mean_interval}
##### 1. Calculating Avg. Steps:
library(dplyr)
library(ggplot2)
Interval<- activity%>%
        group_by(interval)%>%
        filter(!is.na(steps))%>%
        summarise(avg_steps=mean(steps, na.rm=TRUE))
head(Interval)

##### 1.a Plotting Avg. Steps:
ggplot(Interval, aes(x =interval , y=avg_steps)) +
        geom_line(color="purple", size=2) +
        labs(title = "Avergate Daily Steps for each interval", x = "Interval", y = "Average Steps per day")

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
maxsteps<-Interval[which.max(Interval$avg_steps),]
maxsteps
print(paste("The maximum number of steps in a five minute interval was: ", maxsteps$avg_steps))

```

        
**4. Imputing missing values**
```{r missing_value}

#total missing data

missingdata <- sum(is.na(activity$steps))
print(paste("There are", missingdata, "missing data points."))

# Replacing NA value with mean value

activity_1 <- activity
activity_1$steps[is.na(activity_1$steps)] <- mean(activity$steps, na.rm=TRUE)
activity_2 <- aggregate(steps ~ date, data= activity_1, sum, na.rm=TRUE)
hist(activity_2$steps, breaks=40, main="Total Steps per Day \n Adjusted Data",
     xlab="Steps", ylab="Frequency", col='red')


print(paste("median for adjusted data is ", median(activity_2$steps)))
print(paste("mean for adjusted data is ", mean(activity_2$steps)))

```

In conclusion: the impact of imputing missing data with the average number of steps in the same 5-min interval is that both the mean and the median are same : 10766.19

**5. Are there differences in activity patterns between weekdays and weekends?**

```{r week}
activity_1$date <- as.Date(activity_1$date)
activity_1$day_name <- weekdays(activity_1$date)
table(activity_1$day_name)
activity_1$week_ind <- as.factor(ifelse(activity_1$day_name == "星期六" | activity_1$day_name == "星期日", "weekend", "weekday"))
library(lattice)
plot_data <- aggregate(steps ~ interval + week_ind, activity_1, mean)
xyplot(steps ~ interval | (week_ind), data=plot_data, aspect=1/2, type="l")

plot_data_weekday<-subset(plot_data, week_ind == "weekday")
plot_data_weekend<-subset(plot_data, week_ind == "weekend")

print(paste("max step in weekday is ", plot_data_weekday[which.max(plot_data_weekday$steps), ]$steps, "in interval", plot_data[which.max(plot_data_weekday$steps), ]$interval))

print(paste("max step in weekends is ", plot_data_weekend[which.max(plot_data_weekend$steps), ]$steps, "in interval", plot_data[which.max(plot_data_weekend$steps), ]$interval))

```
